---
title: "remlTemplate"
author: "SRE"
date: Sys.Date()
output: html_document
editor_options: 
  chunk_output_type: console
---

# libraries

```{r libraries}
library(EML)
library(tidyverse)
library(readxl)
library(capeml)
library(gioseml)
```

# dataset details

```{r dataset_details}
projectid <- 649
packageIdent <- "knb-lter-cap.649.2"
```

# database connections

```{r connections::postgres::local, eval=FALSE}
# functions and setting for connecting to postgres
source("~/Documents/localSettings/pg_local.R")
```

```{r connections::postgres::prod, eval=FALSE}
# functions and setting for connecting to postgres
source("~/Documents/localSettings/pg_prod.R")
```

```{r connections::mysql::prod, eval=FALSE}
# functions and setting for connecting to mysql
source("~/Documents/localSettings/mysql_prod.R")
```

# dataTable

```{r CR1000-EC, eval=TRUE}

OPEC <- dbGetQuery(pg, '
SELECT
  ec.date_time,
  --ec.record,
  --ec.measurement_id,
  m.observation,
  ec.value
FROM fluxtower.cr1000ec ec
JOIN fluxtower.measurements m ON (ec.measurement_id = m.id)
ORDER BY ec.measurement_id, ec.date_time;')

OPEC <- OPEC %>%
  spread(key = observation, value = value)

write_attributes(OPEC)
# write_factors(OPEC)

OPEC_desc <- "Open Path Eddy Covariance (OPEC) data, inlcuding gas analyzer, 3D sonic anemometer, and temperature and relative humidity sensor data measured at the CAP LTER flux tower located in the west Phoenix, AZ neighborhood of Maryvale. Data included here are 30-minute averages of values collected at 10 Hz."

OPEC_DT <- create_dataTable(dfname = OPEC,
  description = OPEC_desc)

```

```{r CR23X, eval=TRUE}

radiometer_rain <- dbGetQuery(pg, '
SELECT
  cr23x.date_time,
  --ec.record,
  --ec.measurement_id,
  m.observation,
  cr23x.value
FROM fluxtower.cr23x
JOIN fluxtower.measurements m ON (cr23x.measurement_id = m.id)
ORDER BY cr23x.measurement_id, cr23x.date_time;')

radiometer_rain <- radiometer_rain %>%
  spread(key = observation, value = value)

write_attributes(radiometer_rain)
# write_factors(radiometer_rain)

radiometer_rain_desc <- "radiometer and rain gauge (30-minute averages) data measured at the CAP LTER flux tower located in the west Phoenix, AZ neighborhood of Maryvale"

radiometer_rain_DT <- create_dataTable(dfname = radiometer_rain,
  description = radiometer_rain_desc)

```

```{r CR1000-Soil, eval=TRUE}

groundSensors <- dbGetQuery(pg, '
SELECT
  cr1000soil.date_time,
  --ec.record,
  --ec.measurement_id,
  m.observation,
  cr1000soil.value
FROM fluxtower.cr1000soil
JOIN fluxtower.measurements m ON (cr1000soil.measurement_id = m.id)
ORDER BY cr1000soil.measurement_id, cr1000soil.date_time;')

groundSensors <- groundSensors %>%
  spread(key = observation, value = value)

write_attributes(groundSensors)
# write_factors(groundSensors)

groundSensors_desc <- "data (30-minute averages) from 3 soil moisture sensors, 4 soil temperature probes, and 2 heat flux plates buried at multiple depths at the CAP LTER flux tower located in the west Phoenix, AZ neighborhood of Maryvale"

groundSensors_DT <- create_dataTable(dfname = groundSensors,
  description = groundSensors_desc)

```

```{r maint-log, eval=TRUE}

maintenance_log <- read_csv("maintenance_log.csv")

write_attributes(maintenance_log)
# write_factors(maintenance_log)

maintenance_log_desc <- "log of maintenance activity at the CAP LTER flux tower located in the west Phoenix, AZ neighborhood of Maryvale"

maintenance_log_DT <- create_dataTable(dfname = maintenance_log,
  description = maintenance_log_desc)

```


# title

```{r title}

title <- 'Long-term monitoring of micrometeorological conditions at the CAP LTER flux tower located in the west Phoenix, AZ neighborhood of Maryvale, ongoing since 2010'
```

# people

See the gioseml package for examples of creating people resources from scratch.

```{r people}

# creator(s) - required

winston <- create_role(firstName = "w",
                       lastName = "chow",
                       roleType = "creator")

dan <- create_role(firstName = "dan",
                       lastName = "childers",
                       roleType = "creator")

# fabio <- create_role(giosPersonId = 23857,
#                      roleType = "creator")

creators <- list(winston, dan)

# metadata provider - required

winston <- create_role(firstName = "w",
                       lastName = "chow",
                       roleType = "metadata")

tom <- EML::eml$metadataProvider(
  individualName = EML::eml$individualName(
    givenName = "Thomas",
    surName = "Volo"),
  electronicMailAddress = "Thomas.Volo@asu.edu",
  organizationName = "Arizona State University",
  id = "649.metadata_provider.thomas.volo",
)

stevan <- create_role(firstName = "st",
                       lastName = "earl",
                       roleType = "metadata")

metadataProvider <- list(winston, tom, stevan) 

```

# keywords

```{r keywords}

# CAP IRTs for reference (be sure to include these as appropriate):
# https://sustainability.asu.edu/caplter/research/

write_keywords()
```

# methods

Methods are automatically read from a `methods.md` file in the project
directory. If more elaborate methods are required, e.g., to incorporate
provenance, use the enhancedMethods approach outlined below.

Use this extended approach of developing methods if provenance data are
required or there are multiple methods files, otherwise the `create_dateset()`
function will look for a methods.md file in the working directory (or a file
path and name can be passed).

## methods and provenance

```{r methods_provenance, eval=FALSE}

# methods from file tagged as markdown
main <- list(description = list(read_markdown("methods.md")))

library(EDIutils)

# provenance: naip
naip <- emld::as_emld(EDIutils::api_get_provenance_metadata("knb-lter-cap.623.1"))
naip$`@context` <- NULL
naip$`@type` <- NULL

# provenance: lst
landSurfaceTemp <- emld::as_emld(EDIutils::api_get_provenance_metadata("knb-lter-cap.677.1"))
landSurfaceTemp$`@context` <- NULL
landSurfaceTemp$`@type` <- NULL

enhancedMethods <- EML::eml$methods(methodStep = list(main, naip, landSurfaceTemp))

```

# coverages

```{r coverages}

begindate <- "2010-09-01"
enddate <- max(OPEC$date_time)
enddate <- format(enddate, format = "%Y-%m-%d") 
geographicDescription <- "CAP LTER study area"
coverage <- set_coverage(begin = begindate,
                         end = enddate,
                         geographicDescription = geographicDescription,
                         west = -112.1426, east = -112.1426,
                         north = +33.4838, south = +33.4838)

```


# dataset

Optionally, provide: scope, abstract, methods, keywords, publication date.
Projects scopes include lter (default), urex, and som.

```{r construct-dataset}

dataset <- create_dataset()
```

# add dataTable

```{r dataSet$dataTable}

# add dataTables if relevant

print(ls(pattern = "_DT"))

if (length(ls(pattern = "_DT")) > 0) {
  
  listOfDataTables <- lapply(ls(pattern = "_DT"), function(DT) { get(DT) } )
  
  dataset$dataTable  <- listOfDataTables  
  
}

# or add manually
# dataset$dataTable <- list(dataTableOne, dataTableTwo)

```


# customUnits

```{r custom-units, eval=TRUE}

custom_units <- rbind(
  data.frame(id = "wattPerMeterSquared",
             unitType = "irradiance",
             parentSI = "wattPerMeterSquared",
             multiplierToSI = 1,
             description = "radiant flux received by a surface per unit area"),
  data.frame(id = "gramPerMeterCubed",
             unitType = "density",
             parentSI = "kilogramPerCubicMeter",
             multiplierToSI = 0.001,
             description = "mass in grams divided by volume in cubic meters"),
  data.frame(id = "milligramsPerMeterSquaredPerSecond",
             unitType = "arealMassDensityRate",
             parentSI = "kilogramsPerMeterSquaredPerSecond",
             multiplierToSI = 0.000001,
             description = "rate of mass flow per unit area in units of milligrams per meter squared per second"),
  data.frame(id = "gramsPerMeterSquaredPerSecond",
             unitType = "arealMassDensityRate",
             parentSI = "kilogramsPerMeterSquaredPerSecond",
             multiplierToSI = 0.001,
             description = "rate of mass flow per unit area in units of grams per meter squared per second"),
  data.frame(id = "meterSquaredPerSecondSquared",
             unitType = "unknown",
             parentSI = "unknown",
             multiplierToSI = "unknown",
             description = "covariance of N-S (metersPerSecond) and E-W (metersPerSecond) wind speeds"),
  data.frame(id = "meterCelsiusPerSecond",
             unitType = "unknown",
             parentSI = "unknown",
             multiplierToSI = "unknown",
             description = "covariance of virtual temperature (celsius) and wind speed (metersPerSecond)"))

unitList <- set_unitList(custom_units,
                         as_metadata = TRUE)

```

# literature cited

```{r literature cited, eval=FALSE}

# add literature cited if relevant
library(rcrossref)
library(EML)

mccafferty <- cr_cn(dois = "https://doi.org/10.1186/s40317-015-0075-2", format = "bibtex")
mccaffertycit <- EML::eml$citation(id = "https://doi.org/10.1186/s40317-015-0075-2")
mccaffertycit$bibtex <- mccafferty 

citations <- list(
  citation = list(
    mccaffertycit
  ) # close list of citations
) # close citation

dataset$literatureCited <- citations 

```

# eml

```{r construct_eml, eval=TRUE}

eml <- create_eml()
```

```{r write_eml}

# write the eml to file
write_eml(eml, paste0(packageIdent, ".xml"))
```
