---
title: "remlTemplate"
author: "SRE"
date: Sys.Date()
output: html_document
editor_options: 
  chunk_output_type: console
---

# libraries

```{r libraries}
library(tidyverse)
library(readxl)
library(capeml)
library(gioseml)
```

# data entities

## CR1000-EC

```{r CR1000-EC, eval=TRUE}

OPEC <- dbGetQuery(pg, '
SELECT
  ec.date_time,
  --ec.record,
  --ec.measurement_id,
  m.observation,
  ec.value
FROM fluxtower.cr1000ec ec
JOIN fluxtower.measurements m ON (ec.measurement_id = m.id)
ORDER BY ec.measurement_id, ec.date_time;')

OPEC <- OPEC %>%
  spread(key = observation, value = value)

# write_attributes(OPEC)
# write_factors(OPEC)

OPEC_desc <- "Open Path Eddy Covariance (OPEC) data, inlcuding gas analyzer, 3D sonic anemometer, and temperature and relative humidity sensor data measured at the CAP LTER flux tower located in the west Phoenix, AZ neighborhood of Maryvale. Data included here are 30-minute averages of values collected at 10 Hz."

OPEC_DT <- create_dataTable(
  dfname = OPEC,
  description = OPEC_desc,
  dateRangeField = "date_time"
)

```

## CR23X

```{r CR23X, eval=TRUE}

radiometer_rain <- dbGetQuery(pg, '
SELECT
  cr23x.date_time,
  --ec.record,
  --ec.measurement_id,
  m.observation,
  cr23x.value
FROM fluxtower.cr23x
JOIN fluxtower.measurements m ON (cr23x.measurement_id = m.id)
ORDER BY cr23x.measurement_id, cr23x.date_time;')

radiometer_rain <- radiometer_rain %>%
  spread(key = observation, value = value)

# write_attributes(radiometer_rain)
# write_factors(radiometer_rain)

radiometer_rain_desc <- "radiometer and rain gauge (30-minute averages) data measured at the CAP LTER flux tower located in the west Phoenix, AZ neighborhood of Maryvale"

radiometer_rain_DT <- create_dataTable(
  dfname = radiometer_rain,
  description = radiometer_rain_desc,
  dateRangeField = "date_time"
)

```

## CR1000-Soil

```{r CR1000-Soil, eval=TRUE}

groundSensors <- dbGetQuery(pg, '
SELECT
  cr1000soil.date_time,
  --ec.record,
  --ec.measurement_id,
  m.observation,
  cr1000soil.value
FROM fluxtower.cr1000soil
JOIN fluxtower.measurements m ON (cr1000soil.measurement_id = m.id)
ORDER BY cr1000soil.measurement_id, cr1000soil.date_time;')

groundSensors <- groundSensors %>%
  spread(key = observation, value = value)

# write_attributes(groundSensors)
# write_factors(groundSensors)

groundSensors_desc <- "data (30-minute averages) from 3 soil moisture sensors, 4 soil temperature probes, and 2 heat flux plates buried at multiple depths at the CAP LTER flux tower located in the west Phoenix, AZ neighborhood of Maryvale"

groundSensors_DT <- create_dataTable(
  dfname = groundSensors,
  description = groundSensors_desc,
  dateRangeField = "date_time"
)

```

## maintenance log

```{r maint-log, eval=TRUE}

maintenance_log <- read_csv("maintenance_log.csv")

# write_attributes(maintenance_log)
# write_factors(maintenance_log)

maintenance_log_desc <- "log of maintenance activity at the CAP LTER flux tower located in the west Phoenix, AZ neighborhood of Maryvale"

maintenance_log_DT <- create_dataTable(
  dfname = maintenance_log,
  description = maintenance_log_desc,
  dateRangeField = "date"
)

```


# people

See the gioseml package for examples of creating people resources from scratch.

```{r people}

# creator(s) - required

winston <- create_role(
  firstName = "w",
  lastName = "chow",
  roleType = "creator")

dan <- create_role(
  firstName = "dan",
  lastName = "childers",
  roleType = "creator")

creators <- list(winston, dan)

# metadata provider - required

winston <- create_role(
  firstName = "w",
  lastName = "chow",
  roleType = "metadata")

tom <- EML::eml$metadataProvider(
  individualName = EML::eml$individualName(
    givenName = "Thomas",
    surName = "Volo"),
  electronicMailAddress = "Thomas.Volo@asu.edu",
  organizationName = "Arizona State University"
)

stevan <- create_role(
  firstName = "st",
  lastName = "earl",
  roleType = "metadata")

metadataProvider <- list(winston, tom, stevan)

```

# keywords

```{r keywords}

# CAP IRTs for reference (be sure to include these as appropriate):
# https://sustainability.asu.edu/caplter/research/

write_keywords()
```

# methods

Methods are automatically read from a `methods.md` file in the project
directory. If more elaborate methods are required, e.g., to incorporate
provenance, use the enhancedMethods approach outlined below.

Use this extended approach of developing methods if provenance data are
required or there are multiple methods files, otherwise the `create_dateset()`
function will look for a methods.md file in the working directory (or a file
path and name can be passed).


# coverages

```{r coverages}

begindate <- "2010-09-01"

enddate <- max(
  max(as.Date(OPEC$date_time)),
  max(as.Date(groundSensors$date_time)),
  max(as.Date(radiometer_rain$date_time))
)
enddate <- format(enddate, format = "%Y-%m-%d")

geographicDescription <- "CAP LTER study area"

coverage <- EML::set_coverage(
  begin = begindate,
  end = enddate,
  geographicDescription = geographicDescription,
  west = -112.1426,
  east = -112.1426,
  north = +33.4838,
  south = +33.4838
)

```


# dataset

Optionally, provide: scope, abstract, methods, keywords, publication date.
Projects scopes include lter (default), urex, and som.

```{r construct-dataset}

dataset <- create_dataset()
```

# add dataTable

```{r dataSet$dataTable}

# add dataTables if relevant

print(ls(pattern = "_DT"))

if (length(ls(pattern = "_DT")) > 0) {

  listOfDataTables <- lapply(ls(pattern = "_DT"), function(DT) { get(DT) } )

  dataset$dataTable  <- listOfDataTables

}

# or add manually
# dataset$dataTable <- list(dataTableOne, dataTableTwo)

```


# customUnits

```{r custom-units, eval=TRUE}

custom_units <- rbind(
  data.frame(id = "wattPerMeterSquared",
             unitType = "irradiance",
             parentSI = "wattPerMeterSquared",
             multiplierToSI = 1,
             description = "radiant flux received by a surface per unit area"),
  data.frame(id = "gramPerMeterCubed",
             unitType = "density",
             parentSI = "kilogramPerCubicMeter",
             multiplierToSI = 0.001,
             description = "mass in grams divided by volume in cubic meters"),
  data.frame(id = "milligramsPerMeterSquaredPerSecond",
             unitType = "arealMassDensityRate",
             parentSI = "kilogramsPerMeterSquaredPerSecond",
             multiplierToSI = 0.000001,
             description = "rate of mass flow per unit area in units of milligrams per meter squared per second"),
  data.frame(id = "gramsPerMeterSquaredPerSecond",
             unitType = "arealMassDensityRate",
             parentSI = "kilogramsPerMeterSquaredPerSecond",
             multiplierToSI = 0.001,
             description = "rate of mass flow per unit area in units of grams per meter squared per second"),
  data.frame(id = "meterSquaredPerSecondSquared",
             unitType = "unknown",
             parentSI = "unknown",
             multiplierToSI = "unknown",
             description = "covariance of N-S (metersPerSecond) and E-W (metersPerSecond) wind speeds"),
  data.frame(id = "meterCelsiusPerSecond",
             unitType = "unknown",
             parentSI = "unknown",
             multiplierToSI = "unknown",
             description = "covariance of virtual temperature (celsius) and wind speed (metersPerSecond)"))

unitList <- EML::set_unitList(
  custom_units,
  as_metadata = TRUE
)

```


# eml

```{r construct-eml, eval=TRUE}

eml <- capeml::create_eml()
```

```{r validate-eml, eval=TRUE}

EML::eml_validate(eml)
```

```{r eml-to-file, eval=TRUE}

capeml::write_cap_eml()
```


# file placement

## package details

```{r package-details, eval=TRUE}

# retrieve package details from config.yaml

if (!file.exists("config.yaml")) {
  stop("config.yaml not found")
}
packageIdent <- yaml::yaml.load_file("config.yaml")$packageIdent
packageNum <- yaml::yaml.load_file("config.yaml")$packageNum
```

## AWS settings

```{r S3_helper_functions}

# functions and setting for uploading to S3

library(aws.s3)
source("~/Documents/localSettings/aws.s3")
```

## upload data files

```{r preview_data_file_to_upload}

# preview data set files that will be uploaded to S3
list.files(pattern = paste0(packageNum, "_"))
```

```{r upload_data_S3}

# upload files to S3

lapply(list.files(pattern = paste0(packageNum, "_")), data_to_amz)
```

## copy and move XML as needed

```{r duplicate_xml_as_needed}

# XML to S3
eml_to_amz(list.files(pattern = paste0(packageIdent, ".xml")))

# XML to cap-data-eml
file.copy(
  from = list.files(pattern = paste0(packageIdent, ".xml")),
  to = "/home/srearl/localRepos/cap-metadata/cap-data-eml/"
)
```

## optionally delete data and xml files

```{r optional: preview_files_to_delete}

# preview files to delete
dir(pattern = paste0(packageNum, "_"))
list.files(pattern = paste0(packageIdent, ".xml"))
```

```{r optional: delete_files}

# delete files
file.remove(dir(pattern = paste0(packageNum, "_")))
file.remove(list.files(pattern = paste0(packageIdent, ".xml")))
```
